Constraining gravitational wave amplitude birefringence with GWTC-3

### 2022-08-10 src/tex/ms.tex
Update ms.tex

Since Einstein proposed his theory of general relativity (GR), it was tested in a wide range of length scales.
After a century, we know that GR do not agree with quantum theories at some length scale.
To unify both theories, we need to study the possibility of different beyond-GR theories.
Some beyond-GR theories such as Chern-Simons theory (CS) suggested that there is gravitational wave (GW) amplitude birefringence,
while GR predicts there is no birefringence.
GW amplitude birefringence is a property of space which consists of the enhancement of one GW polarisation over the other during the propagation of the wave.

% To illustrate the effect of amplitude birefringence,
% the amplitudes of two polarisations of a artificial GW signal are shown in figure \textbf{[X]}.

@@

\begin{figure*}[h]
    \script{GW150914_corner.py}
    \includegraphics[width=\textwidth]{figures/GW150914_corner.png}
    \caption{The posterior of $\kappa$, luminosity distance $d_L$ and $\cos{\iota}$ for GW150914.
    The three sets of plots are the PEs done by LIGO with GR \citep{LIGO_2021},
    the PE done by us with the frequency independent birefringence and the frequency dependent birefringence respectively.
    Note that there is no posterior of $\kappa$ for the PE from LIGO, as the LIGO PE is based on GR,
    which does not suggest GW amplitude birefringence.}
    \label{fig:GW150914_corner}
\end{figure*}

@@

@@

\begin{enumerate}
    \item Introduction
    \begin{enumerate}
        \item Motivation
        \begin{enumerate}
            \item GR broke down at some scale, find a new theory
            \item CS: beyond-GR theory
            \item test CS property (GW amplitude birefringence)
        \end{enumerate}
        \item Amplitude Birefringence
        \begin{enumerate}
            \item GW amplitude birefringence is a property of space
            \item birefringence: enhance one and suppress another
            \item figure to illustrate the effect of birefringence
        \end{enumerate}
        \item Amplitude Ratio
        \begin{enumerate}
            \item GR: amplitude ratio depends on $\iota$ only
            \item nonGR: also depends on $\kappa$, $d_C$ and $f$
        \end{enumerate}
    \end{enumerate}
    \item Method
    \begin{enumerate}
        \item PE on GWTC-3 to constrain $\kappa$
        \item Assume GW are generated as GR described
        \item birefringence is considered during the propagation
        \item waveform modification
    \end{enumerate}
    \item Results
    \begin{enumerate}
        \item GW150914
        \begin{enumerate}
            \item figure with frequency-dependent and frequency-independent
            \item posterior from $\cos\iota$ looks different
            \item frequency-dependent: degeneracy between $\iota$ and $\kappa$
            \item frequency-independent: degeneracy broken and $\kappa$ near $0$
        \end{enumerate}
        \item GWTC-3
        \begin{enumerate}
            \item figure of $\kappa$ of all events
            \item constraint on $\kappa$
            \item improvement of constraint on $\kappa$ compare to previous studies
        \end{enumerate}
    \end{enumerate}
    \item Discussion
    \begin{enumerate}
        \item Limitation
        \begin{enumerate}
            \item orientation of detector (polarisations)
            \item sensitivity of detector
        \end{enumerate}
        \item Future Study
        \begin{enumerate}
            \item More events from detector
            \item Higher order terms
        \end{enumerate}
    \end{enumerate}
\end{enumerate}

>>>>

Since Einstein proposed his theory of general relativity (GR), it was tested in a wide range of length scales.
After a century, we know that GR do not agree with quantum theories at some length scale.
To unify both theories, we need to study the possibility of different beyond-GR theories.
Some beyond-GR theories such as Chern-Simons theory (CS) suggested that there is gravitational wave (GW) amplitude birefringence,
while GR predicts there is no birefringence.

GW amplitude birefringence is a property of space which consists of the enhancement of one GW polarisation over the other during the propagation of the wave.
The longer the GW travels, the larger the effect of birefringence.
To illustrate the effect of amplitude birefringence,
the observed amplitudes of two polarisations of an artificial GW signal example are shown in figure \textbf{[X]}.
\textbf{[add figure similar to the animation in the website]}
With birefringence, the amplitude is modified during the propagation,
and the observed amplitudes will be different compare to the amplitudes when they are generated at the binary.

@@

\begin{figure}[h!]
    \script{GW150914_corner.py}
    \includegraphics[width=\columnwidth]{figures/GW150914_corner.png}
    \caption{The posterior of $\kappa$, luminosity distance $d_L$ and $\cos{\iota}$ for GW150914.
    The three sets of plots are the PEs done by LIGO with GR \citep{LIGO_2021},
    the PE done by us with the frequency independent birefringence and the frequency dependent birefringence respectively.
    Note that there is no posterior of $\kappa$ for the PE from LIGO, as the LIGO PE is based on GR,
    which does not suggest GW amplitude birefringence.}
    \label{fig:GW150914_corner}
\end{figure}

@@

We then perform PE on $71$ binary black hole merger events in GWTC-3 to further constrain $\kappa$.
The posteriors for $\kappa$ of the events are shown in figure \textbf{[X]}, which shows $\kappa$ tend to be $0$.
The combined posterior for $\kappa$ was calculated and shown in figure \textbf{[X]}, which allows us to constrain $\kappa$ within \textbf{[the number we get]}.

@@

\begin{enumerate}
    \item Introduction
    \begin{enumerate}
        \item Motivation
        \begin{enumerate}
            \item GR broke down at some scale, find a new theory
            \item CS: beyond-GR theory
            \item test CS property (GW amplitude birefringence)
        \end{enumerate}
        \item Amplitude Birefringence
        \begin{enumerate}
            \item GW amplitude birefringence is a property of space
            \item birefringence: enhance one and suppress another
            \item figure to illustrate the effect of birefringence
        \end{enumerate}
        \item Amplitude Ratio
        \begin{enumerate}
            \item GR: amplitude ratio depends on $\iota$ only
            \item nonGR: also depends on $\kappa$, $d_C$ and $f$
        \end{enumerate}
    \end{enumerate}
    \item Method
    \begin{enumerate}
        \item Parameter Estimation
        \begin{enumerate}
            \item PE on GWTC-3 to constrain $\kappa$
            \item Assume GW are generated as GR described
            \item birefringence is considered during the propagation
            \item waveform modification
        \end{enumerate}
    \end{enumerate}
    \item Results
    \begin{enumerate}
        \item GW150914
        \begin{enumerate}
            \item figure with frequency-dependent and frequency-independent
            \item posterior from $\cos\iota$ looks different
            \item frequency-dependent: degeneracy between $\iota$ and $\kappa$
            \item frequency-independent: degeneracy broken and $\kappa$ near $0$
        \end{enumerate}
        \item GWTC-3
        \begin{enumerate}
            \item figure of $\kappa$ of all events
            \item constraint on $\kappa$
            \item improvement of constraint on $\kappa$ compare to previous studies
        \end{enumerate}
    \end{enumerate}
    \item Discussion
    \begin{enumerate}
        \item Limitation
        \begin{enumerate}
            \item orientation of detector (polarisations)
            \item sensitivity of detector
        \end{enumerate}
        \item Future Study
        \begin{enumerate}
            \item More events from detector
            \item Higher order terms
        \end{enumerate}
    \end{enumerate}
\end{enumerate}
